Embracing the Laws of Physics: \\ Three Reversible Models of Computation

### 2018-06-25 survey/survey.tex
remove comment out comment
commentary

% \jc{While I really like this model, if we are going to keep it, we
% really need a much beefier introduction. This is the first mention of
% resources (and events and time). So we really should preface this by
% saying that, once we have $\Pi$, there is another \emph{completely
% different} but compatible interpretation in terms of resources. But
% we should also think of whether this fits. I actually think that it
% does fit in this paper, but perhaps not in this section. Maybe
% also move to section 5?}

The textual representation of $\Pi$ is quite cumbersome: the
underlying principles however become quite intuitive when represented
in a more suggestive visual notation. There exist various graphical
notations for the monoidal categories that model $\Pi$ and related
languages including Penrose diagrams~\cite{selinger-graphical} and
Geometry of Interaction (GoI) machines and Proof
Nets~\cite{Mackie2011,DBLP:conf/popl/Mackie95}. In this section, we
propose a novel graphical representation of $\Pi$ types and
combinators drawn from the resource usage embedded in the operational
semantics. As suggested by the operational semantics, the type $A * B$
can be viewed as denoting \emph{spatial composition}, where
\emph{both} $A$ and $B$ must happen in \emph{some} order or perhaps
simultaneously, and where $A + B$ denotes \emph{temporal composition},
where exactly \emph{one of} $A$ and $B$ must happen at a given time.

>>>>

The textual representation of $\Pi$ is quite cumbersome: the
underlying principles however become quite intuitive when represented
in a more suggestive visual notation. There exist various graphical
notations for the monoidal categories that model $\Pi$ and related
languages including Penrose diagrams~\cite{selinger-graphical} and
Geometry of Interaction (GoI) machines and Proof
Nets~\cite{Mackie2011,DBLP:conf/popl/Mackie95}. In this section, we
propose a novel graphical representation of $\Pi$ types and
combinators drawn from the resource usage embedded in the operational
semantics. As suggested by the operational semantics, the type $A * B$
can be viewed as denoting \emph{spatial composition}, where
\emph{both} $A$ and $B$ must happen in \emph{some} order or perhaps
simultaneously, and where $A + B$ denotes \emph{temporal composition},
where exactly \emph{one of} $A$ and $B$ must happen at a given time.
\jc{I think the temporal aspect is stretching things too far. Could we
call it ``choice composition''? (Or something better). To me spatial
composition somehow seems like sequential composition (i.e. composition
in time), which is definitely not what we want. I think the informal
description (both although perhaps concurrently, and exactly one of)
is better. In section 3 we called them sequential, choice and
parallel. We should be consistent.}

\jc{An additional ``problem'' is that since our circuits have a
left-to-right orientation, $A * B$ drawn as boxes left-to-right
gives an implicit ordering that is going to be hard to ignore.
I'm starting to think that the diagrams in the MO answer might
be better: use a colour (or like in Selinger's paper, at the
very end), symbols in the nodes that split.  So both would
be displayed as vertical stacks (only). Because I think that
swap* is too hard to justify in this notation. Of course,
if we could do things in 3D and $*$ was displayed using
(say) depth, that would be different again.}
